### 2016-06-29 reasoning.tex
edits

The rule \rulelabel{RG-Var} defines the rely-guarantee judgment for
shared variable reads inside a transaction $T_i$. It requires $\I$ to
be stable, and pre- and post- conditions to be stable relative to
$\I$.  The quantified premise effectively requires a proof that if the
abstract machine of Fig.~\ref{fig:txnimp} takes a step starting from
an execution $\E$ that satisfies the precondition $P$, then the
resultant execution $\E'$ satisfies the postcondition $Q$, and that
the guarantee $G$ faithfully captures the transition from $\E$ to
$\E'$. It is informative to compare this premise with the premise of
the \rulelabel{E-Aux} reduction rule of Fig.~\ref{fig:txnimp}. Similar
premises also appear in \rulelabel{RG-Asgn} and \rulelabel{RG-Txn}
rules, which define rely-guarantee judgments for assignments and
transactions, respectively. \rulelabel{RG-Var} however also requires
the return value ($\interp{S}(X)$) of the read to satisfy the
assertion $C$ meant for the value. \rulelabel{RG-Arith} defines the RG
judgment for an arithmetic expression $e_1\pm e_2$ in terms of the
corresponding judgments for the constitutent expressions $e_1$ and
$e_2$. The quantified premise requires any value resulting from
evaluating $e_1 \pm e_2$ to satisfy the assertion $C$, provided that
$e_1$ and $e_2$ always evaluate to values that satisfy $C_1$ and
$C_2$, respectively. The rules for sequential and parallel composition
of commands are essentially the same as their counterparts in a
standard rely-guarantee formulation and hence elided.

@@

The \rulelabel{RG-Conseq} rule define ways to strengthen or weaken
relations and assertions associated with the RG judgment of
transaction-bound expressions. Similar rules exist for
transaction-bound and top-level commands, but are not
shown.\footnote{The supplementary materials provide the complete set
  of rules.} As is the case with a standard rely-guarantee
formulation,the rules allow the precondition $P$ and the rely relation
$R$ to be strengthened\footnote{A stronger relation is a smaller set
  than a weaker relation, hence a stronger $R$ admits fewer
  interferences than a weaker one}, and the postcondition $Q$ (also,
$C$ in the case of expressions) and the guarantee relation $G$ to be
weakened.  The most notable aspect of the \rulelabel{RG-Conseq} rules
is that they allow the trace invariant $\I$ to be
strengthened. Considering that $\I$ captures isolation properties,
this means that a program proven correct under weaker isolation levels
is also correct under stronger ones.  Parametricity over the trace
invariant $\I$, combined with the ability to strengthen $\I$ as
needed, allows our proof system to support a highly flexible proof
strategy to prove programs correct over various isolation
variants. For example, programmers can \emph{define} isolation
guarantees of their choice (by defining $\I$ appropriately) and then
prove programs correct assuming the guarantees hold.  The soundness of
strengthening $\I$ assures that the program can now be executed safely
on any data store that offers isolation guarantees at least as strong
as those assumed.

@@

\begin{definition}[\bfseries Step-indexed reflexive transitive closure]
Forall $A:\text{Type}$, $R: A \rightarrow A \rightarrow \mathbb{P}$, and $n :
\mathbb{N}$, step-indexed reflexive transitive closure $R^n$ of $R$ is
the smallest relation satisfying the following
properties:
\begin{itemize}
\item $\forall (x:A).\, R^0 (x,x)$
\item $\forall (n:\mathbb{N})(x,y,z : A).\, R(x,y) \conj R^n(y,z) \Rightarrow
R^{n+1}(x,z)$
\end{itemize}
\end{definition}

\begin{definition}[\bfseries Interleaved step and multi-step relations]
An Interleaved step relation interleaves thread-local reductions with
interference from concurrent threads captured as the rely relation
($R$).  It is defined thus:
\begin{smathpar}
\begin{array}{lcl}
\I \vdash (c,\E) \rstepsto (c',\E') & \defeq & \I \vdash 
  (c,\E) \stepsto (c',\E') \\
  &   & \disj (c' = c \conj R(\E, \E') \conj \I(\E'))\\
\end{array}
\end{smathpar}
The interleaved step relation for transaction bound expressions
($\txnbox{e}_i$) and commands ($\txnbox{c}_i$) is defined similarly.
An Interleaved multi-step relation ($\stepssto{n}$) is the step-indexed
reflexive transitive closure of the interleaved step relation.
\end{definition}

\begin{definition}[\bfseries Semantics of the RG judgment]
\label{def:rg-semantics}
The semantics of the RG sextuple $\rg{\mathbb{I},P,R}{c}{G,Q}$ is defined
in terms of the interleaved step relation thus:
\begin{smathpar}
\begin{array}{l}
\hspace*{-0.3in}
\rg{\mathbb{I},P,R}{c}{G,Q} \;\defeq\; \forall \E.\, P(\E)
  \wedge \mathbb{I}(\E) \\
\hspace*{0.4in}\Rightarrow (\forall n,\E'.\; \I \vdash (c,\E) 
    \rstepssto{n} (\cskip,\E') \Rightarrow Q(\E')) \\
\hspace*{0.5in}\conj \texttt{step-guaranteed}(\I,R,G,c,\E)\\
\end{array}
\end{smathpar}
The first conjunct in the consequent is called the \emph{Hoare
consequent} since it ascribes Hoare triple semantics to an RG sextuple.
The second conjunct, called the \emph{guarantee consequent}, uses the
$\texttt{step-guaranteed}$ predicate defined below:
\begin{smathpar}
\begin{array}{l}
\texttt{step-guaranteed}(\I,R,G,c,\E) \;\defeq\; \forall n,\E',c'',\E''.\\
\hspace*{0.2in}\I \vdash (c,\E) \rstepssto{n} (c',\E') \conj \I \vdash (c',\E') \stepsto
  (c'',\E'') \Rightarrow G(\E',\E'')\\
\end{array}
\end{smathpar}
The guarantee consequent requires $G$ to capture the trace effect of
every small-step of $c$, where the reduction can be interleaved by the
interference ($R$) from concurrent threads. The semantics of the RG
sextuple for transaction-bound commands ($\txnbox{c}_i$) is defined
similarly. Expressions, unlike commands, evaluate to a value $v$, and
the semantics of their RG septuple ($\rg{\I,P,R}{\txnbox{e}_i}{G,C,Q}$) differs slightly in that its
Hoare consequent requires the value $v$ to satisfy the assertion $C$. 
\end{definition}

@@

\begin{theorem}[\bfseries Soundness] 
The rely-guarantee judgments defined by the rules in
Fig.~\ref{fig:rg-rules} are sound with respect to
the semantics of Definition~\ref{def:rg-semantics}.\footnote{Full proofs are provided in the supplementary material.}
\end{theorem}
\noindent In particular, if $\rg{\I,P,R}{c}{G,Q}$ can be derived using
the rules of Fig.~\ref{fig:rg-rules}, then (a) every interleaved
multi-step reduction of $c$ starting from a trace that satisfies $P$
and $\I$, results in a trace that satisfies $Q$, and (b) the effect
that every small-step of $c$ has on the trace is contained in $G$.
Soundness of the RG judgment for transaction-bound commands
($\txnbox{c}_i$) is stated similarly.  For expressions, soundness of
the judgment $\rg{\I,P,R}{\txnbox{e}_i}{G,C,Q}$ also proves that $e$
is always evaluated to a value that satisfies $C$.

>>>>

\noindent The rule \rulelabel{RG-Var} defines the rely-guarantee judgment for
shared variable reads inside a transaction $T_i$. It requires $\I$ to
be stable, and pre- and post- conditions to be stable relative to
$\I$.  The quantified premise effectively requires a proof that if the
abstract machine of Fig.~\ref{fig:txnimp} takes a step starting from
an execution $\E$ that satisfies the precondition $P$, then the
resultant execution $\E'$ satisfies the postcondition $Q$, and that
the guarantee $G$ faithfully captures the transition from $\E$ to
$\E'$. It is informative to compare this premise with the premise of
the \rulelabel{E-Aux} reduction rule of Fig.~\ref{fig:txnimp}. Similar
premises also appear in \rulelabel{RG-Asgn} and \rulelabel{RG-Txn}
rules, which define rely-guarantee judgments for assignments and
transactions, respectively. \rulelabel{RG-Var} however also requires
the return value ($\interp{S}(X)$) of the read to satisfy the
assertion $C$ meant for the value. \rulelabel{RG-Arith} defines the RG
judgment for an arithmetic expression $e_1\pm e_2$ in terms of the
corresponding judgments for the constitutent expressions $e_1$ and
$e_2$. The quantified premise requires any value resulting from
evaluating $e_1 \pm e_2$ to satisfy the assertion $C$, provided that
$e_1$ and $e_2$ always evaluate to values that satisfy $C_1$ and
$C_2$, respectively. The rules for sequential and parallel composition
of commands are essentially the same as their counterparts in a
standard rely-guarantee formulation and hence elided.

@@

The \rulelabel{RG-Conseq} rule define ways to strengthen or
weaken relations and assertions associated with the RG judgment of
transaction-bound expressions. Similar rules exist for
transaction-bound and top-level commands, but are not
shown.\footnote{The supplementary materials provide the complete set
  of rules.} As is the case with a standard rely-guarantee
formulation,the rules allow the precondition $P$ and the rely relation
$R$ to be strengthened, and the postcondition $Q$ (also, $C$ in the
case of expressions) and the guarantee relation $G$ to be weakened.
The most notable aspect of the \rulelabel{RG-Conseq} rules is that
they allow the trace invariant $\I$ to be strengthened. Considering
that $\I$ captures isolation properties, this means that a program
proven correct under weaker isolation levels is also correct under
stronger ones.  Parametricity over the trace invariant $\I$, combined
with the ability to strengthen $\I$ as needed, allows our proof system
to support a highly flexible proof strategy to prove programs correct
over various isolation variants. For example, programmers can
\emph{define} isolation guarantees of their choice (by defining $\I$
appropriately) and then prove programs correct assuming the guarantees
hold.  The soundness of strengthening $\I$ ensures that a program can
be executed safely on any data store that offers isolation guarantees
at least as strong as those assumed.

@@

\begin{definition}[\bfseries Step-indexed reflexive transitive closure]
For all $A:\text{Type}$, $R: A \rightarrow A \rightarrow \mathbb{P}$, and $n :
\mathbb{N}$, the step-indexed reflexive transitive closure $R^n$ of $R$ is
the smallest relation satisfying the following
properties:
\begin{itemize}
\item $\forall (x:A).\, R^0 (x,x)$
\item $\forall (n:\mathbb{N})(x,y,z : A).\, R(x,y) \conj R^n(y,z) \Rightarrow
R^{n+1}(x,z)$
\end{itemize}
\end{definition}

\begin{definition}[\bfseries Interleaved step and multi-step relations]
An interleaved step relation interleaves thread-local reductions with
interference from concurrent threads captured as the rely relation
($R$).  It is defined thus:

\begin{smathpar}
\begin{array}{lcl}
\I \vdash (c,\E) \rstepsto (c',\E') & \defeq & \I \vdash 
  (c,\E) \stepsto (c',\E') \\
  &   & \disj (c' = c \conj R(\E, \E') \conj \I(\E'))\\
\end{array}
\end{smathpar}

The interleaved step relation for transaction bound expressions
($\txnbox{e}_i$) and commands ($\txnbox{c}_i$) is defined similarly.
An interleaved multi-step relation ($\stepssto{n}$) is the step-indexed
reflexive transitive closure of the interleaved step relation.
\end{definition}

\begin{definition}[\bfseries Semantics of the RG judgment]
\label{def:rg-semantics}
The semantics of the RG sextuple $\rg{\mathbb{I},P,R}{c}{G,Q}$ is defined
in terms of the interleaved step relation thus:

\begin{smathpar}
\begin{array}{l}
\hspace*{-0.3in}
\rg{\mathbb{I},P,R}{c}{G,Q} \;\defeq\; \forall \E.\, P(\E)
  \wedge \mathbb{I}(\E) \\
\hspace*{0.4in}\Rightarrow (\forall n,\E'.\; \I \vdash (c,\E) 
    \rstepssto{n} (\cskip,\E') \Rightarrow Q(\E')) \\
\hspace*{0.5in}\conj \texttt{step-guaranteed}(\I,R,G,c,\E)\\
\end{array}
\end{smathpar}

\noindent The first conjunct in the consequent is called the \emph{Hoare
consequent} since it ascribes Hoare triple semantics to an RG sextuple.
The second conjunct, called the \emph{guarantee consequent}, uses the
$\texttt{step-guaranteed}$ predicate defined below:

\begin{smathpar}
\begin{array}{l}
\texttt{step-guaranteed}(\I,R,G,c,\E) \;\defeq\; \forall n,\E',c'',\E''.\\
\hspace*{0.2in}\I \vdash (c,\E) \rstepssto{n} (c',\E') \conj \I \vdash (c',\E') \stepsto
  (c'',\E'') \Rightarrow G(\E',\E'')\\
\end{array}
\end{smathpar}

\noindent The guarantee consequent requires $G$ to capture the trace effect of
every small-step of $c$, where the reduction can be interleaved by the
interference ($R$) from concurrent threads. The semantics of the RG
sextuple for transaction-bound commands ($\txnbox{c}_i$) is defined
similarly. Expressions, unlike commands, evaluate to a value $v$, and
the semantics of their RG septuple ($\rg{\I,P,R}{\txnbox{e}_i}{G,C,Q}$) differs slightly in that its
Hoare consequent requires the value $v$ to satisfy the assertion $C$. 
\end{definition}

@@

\begin{theorem}[\bfseries Soundness] 
The rely-guarantee judgments defined by the rules in
Fig.~\ref{fig:rg-rules} are sound with respect to
the semantics of Definition~\ref{def:rg-semantics}.\footnote{Full proofs are provided in the supplementary material.}
\end{theorem}

\noindent In particular, if $\rg{\I,P,R}{c}{G,Q}$ can be derived using
the rules of Fig.~\ref{fig:rg-rules}, then (a) every interleaved
multi-step reduction of $c$ starting from a trace that satisfies $P$
and $\I$, results in a trace that satisfies $Q$, and (b) the effect
that every small-step of $c$ has on the trace is contained in $G$.
Soundness of the RG judgment for transaction-bound commands
($\txnbox{c}_i$) is stated similarly.  For expressions, soundness of
the judgment $\rg{\I,P,R}{\txnbox{e}_i}{G,C,Q}$ also proves that $e$
is always evaluated to a value that satisfies $C$.

### 2016-06-29 related-work.tex
edits

\section{Related Work}

\paragraph{Specifications of weak isolation}
Adya~\cite{adyaphd} specified several isolation levels in terms of
\emph{dependency graphs} between transactions, by specifying what
dependency cycles are forbidden in each case. The operational nature
of Adya's specifications make them suitable for runtime monitoring and
anamoly detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas
declarative nature of our specifications make them suitable for
reasoning about programs. Sivaramakrishnan et al.~\cite{pldi15} specify
isolation levels declaratively as trace well-formedness conditions,
but their specifications implicitly assume a complete trace with only
committed transactions, hence not suitable to operationally reason
about a program as it builds the trace. For instance, their {\sc rc}
specification allows the effects of a not-yet-committed transaction to
be visible to an {\sc rc} transaction.  Nonetheless, our
specifications belong to the same fragment of first-order logic as
theirs, which is known to be decidable. Cerone et
al.~\cite{gotsmanconcur15} present a framework to specify isolation
levels with atomic visibility, but their specifications are also for
the complete traces.  Moreover, their specifications for {\sc si} and
{\sc ser} implicitly assume homogenity, i.e., they require all
transactions to execute at {\sc si} (resp.  {\sc ser}), which is often
not the case in practice. Like ours, all the aforementioned
specification frameworks, use the vocabulary introduced
in~\cite{burckhardt14} to specify replicated data types and forms of
eventual consistency. However, none of them come with an associated
reasoning framework that can use such specifications productively
towards verifying programs under weak isolation.

\paragraph{Reasoning under weak isolation} In~\cite{feketessi}, Fekete
et al. propose a theory to characterize non-serializable executions
that arise under {\sc si}. Fekete~\cite{fekete2005} also proposes an
algorithm that allocates either {\sc si} or {\sc ser} isolation levels
to transactions while guaranteeing the serializability of the
execution. In~\cite{gotsmanpodc16}, Cerone et al. improve on Adya's
{\sc si} specification and use it to derive a static analysis that
determines the safety of substituting {\sc si} with a weaker variant
called \iso{Parallel Snapshot Isolation}~\cite{psi}. All of them focus
on establishing the equivalence of executions between a pair of
isolation levels, without taking application invariants into account.
Bernstein et al.~\cite{bern2000} propose informal semantic conditions
to ensure the satisfaction of application invariants under weaker
isolation levels.  However, no formal proof system has been proposed.
Furthermore, all the aforementioned approaches are tailor-made for a
finite set of well-understood isolation levels (rooted
in~\cite{berenson}) under a pre-defined consistency model (usually
{\sc sc}).

\paragraph{Reasoning under weak consistency} There have been many
recent proposals focusing on reasoning about programs under weak
consistency~\cite{bailisvldb, alvarocalm, gotsmanpopl16,redblueatc,
redblueosdi, ecinec}. All of them assume a system model that offers a
choice between a \emph{coordination-free} weak consistency level ({\sc
ec} in~\cite{redblueosdi, redblueatc, ecinec, alvarocalm, bailisvldb};
{\sc cc} in~\cite{gotsmanpopl16}) and strong consistency ({\sc sc}),
which requires coordination.  They propose various conditions (\eg
\emph{invariant confluence}~\cite{bailisvldb}, \emph{logical
monotonicity}~\cite{alvarocalm}, \emph{invariant
safety}~\cite{redblueatc}), or proof systems (\eg~rely-guarantee
framework of~\cite{gotsmanpopl16}) to show that an allotment of atomic
and fully isolated operations to these levels preserves application
invariants.  In contrast, we admit weakly isolated transactions (sets
of atomic operations), and our system model accepts
\emph{specifications} of consistency and isolation levels drawn from
an expressive logic. Gotsman et al.~\cite{gotsmanpopl16} adapt
\iso{Parallel Snapshot Isolation} to the aforementioned setting by
interpreting it as a consistency level that serializes writes to
objects. They subsequently propose a dedicated proof rule to prove
program invaraints under this model. By parameterizing our proof
system on isolation specifications, we avoid the need to define a
separate proof rule for each new isolation level.

\paragraph{Reasoning under relaxed memory} Considering the
correspondences between axiomatic models of relaxed
memory~\cite{battycpp} and weakly consistent
replication~\cite{burckhardt14}, it is reasonable to expect similar
reasoning approaches to work in both the cases. Ridge~\cite{rgtso}
generalizes rely-guarantee reasoning to x86-TSO relaxed memory model.
Likewise, Vafeadis et al.~\cite{rsl13} generalize concurrent
separation logic (CSL)~\cite{csl} to C11 relaxed memory model.
Ferreira et al.~\cite{ferreira10} propose parameterized operational
semantics for relaxed memory models, but the parameterization is over
a relation between relaxed memory programs and related {\sc sc}
programs. They also prove the soundness of CSL for the weakest memory
model expressible in their semantics, but neither a parameterized (in
the aforementioned sense) CSL, nor a CSL specialized for any relaxed
memory model is proposed. In fact, to the best of our knowledge, there
does not exist a proof system that combines axiomatic semantics of
memory models and operational semantics of programs in a unified
reasoning framework.

>>>>

\section{Related Work}
\label{sec:relatedwork}

\paragraph{Specifying weak isolation.}
Adya~\cite{adyaphd} specified several weak isolation levels in terms
of \emph{dependency graphs} between transactions, by defining the
kinds of dependency cycles that are forbidden in each case. The
operational nature of Adya's specifications make them suitable for
runtime monitoring and anamoly
detection~\cite{kemmevldb,feketesigmod08,pssi2011}, whereas the
declarative nature of our specifications make them suitable for
reasoning about program behaviour. Sivaramakrishnan \emph{et
  al.}~\cite{pldi15} specify isolation levels declaratively as trace
well-formedness conditions, but their specifications implicitly assume
a complete trace with only committed transactions, and is thus not
suitable to reason operationally about a program as it builds the
trace. Cerone \emph{et al.}~\cite{gotsmanconcur15} present a framework
to specify isolation levels with atomic visibility, but their
specifications are also for complete traces; Moreover, their
formulation implicitly assumes homogenity, i.e., they require all
transactions to execute using either {\sc si} or {\sc ser}, which is
often not the case in practice. While all the aforementioned
specification frameworks use the vocabulary introduced
in~\cite{burckhardt14} to specify replicated data types and different
forms of weak consistency, in contrast to this work, none of them come
with an associated reasoning framework that can use such
specifications productively towards verifying programs under weak
isolation.

\paragraph{Reasoning under weak isolation} In~\cite{feketessi}, Fekete
\emph{et al.} propose a theory to characterize non-serializable executions
that arise under {\sc si}. Fekete~\cite{fekete2005} also proposes an
algorithm that allocates either {\sc si} or {\sc ser} isolation levels
to transactions while guaranteeing the serializability of the
execution. In~\cite{gotsmanpodc16}, Cerone et al. improve on Adya's
{\sc si} specification and use it to derive a static analysis that
determines the safety of substituting {\sc si} with a weaker variant
called \iso{Parallel Snapshot Isolation}~\cite{psi}. All of them focus
on establishing the equivalence of executions between a pair of
isolation levels, without taking application invariants into account.
Bernstein et al.~\cite{bern2000} propose informal semantic conditions
to ensure the satisfaction of application invariants under weaker
isolation levels.  However, no formal proof system has been proposed.
Furthermore, all the aforementioned approaches are tailor-made for a
finite set of well-understood isolation levels (rooted
in~\cite{berenson}) under a pre-defined consistency model (usually
{\sc sc}).

\paragraph{Reasoning under weak consistency} There have been several
recent proposals on reasoning techniques for programs executing under
weak consistency~\cite{bailisvldb, alvarocalm,
  gotsmanpopl16,redblueatc, redblueosdi, ecinec}. All of them assume a
system model that offers a choice between a \emph{coordination-free}
weak consistency level (\emph{e.g.}, {\sc ec}~\cite{redblueosdi,
  redblueatc, ecinec, alvarocalm, bailisvldb} or causal consistency
({\sc cc})~\cite{gotsmanpopl16}). In all these efforts, the key
technical issue involves proving that atomic and fully isolated
operations preserve application invariants when executed on these
consistency levels.  In contrast, we admit weakly isolated
transactions (sets of atomic operations), and our system model accepts
\emph{specifications} of consistency and isolation levels drawn from
an expressive logic.  In recent work~\cite{gotsmanpopl16},
\iso{Parallel Snapshot Isolation} is adapted to the aforementioned
setting by interpreting it as a consistency level that serializes
writes to objects; a dedicated proof rule is developed to help prove
prove program invariants hold under this model. By parameterizing our
proof system over a gamut of weak isolation specifications, we avoid
the need to define a separate proof rule for each new isolation level.

\paragraph{Reasoning under relaxed memory} Considering the
correspondences between axiomatic models of relaxed
memory~\cite{battycpp} and weakly consistent
replication~\cite{burckhardt14}, it is reasonable to expect similar
reasoning approaches to work in both the cases. Ridge~\cite{rgtso}
generalizes rely-guarantee reasoning to the x86-TSO memory model.
Likewise, Vafeadis \emph{et al.}~\cite{rsl13} generalize concurrent
separation logic (CSL)~\cite{csl} to the C11 relaxed memory model.
Ferreira et al.~\cite{ferreira10} propose a parameterized operational
semantics for relaxed memory models, but the parameterization is over
a relation between relaxed memory programs and related {\sc sc}
programs. They also prove the soundness of CSL for the weakest memory
model expressible in their semantics, but neither a parameterized (in
the aforementioned sense) CSL, nor a CSL specialized for any relaxed
memory model is proposed. In fact, to the best of our knowledge, there
does not exist a proof system that combines axiomatic semantics of
memory models and operational semantics of programs in a unified
reasoning framework.  \SJ{cite plan b: a buffered memory model for Java.}

### 2016-06-29 store-consistency.tex
edits

A strongly-consistent {\sc sc} store guarantees a total order on all
operations w.r.t $\visZ$ consistent with their chronological order. A
straightforward $\I_s$ for this store is the \C{SC} property
formalized below:
\begin{smathpar}
\begin{array}{l}
  \C{sc}(\E) \;=\; \forall\eta_1,\eta_2.~\{\eta_1,\eta_2\}
  \subseteq \E.\A \conj \id(\eta_1) < \id(\eta_2) \\
  \hspace*{2in}\Rightarrow \underE{\eta_1 \visar \eta_2}
\end{array}
\end{smathpar}
Unfortunately, $\I_s=\C{SC}$ conflicts with all isolation
specifications of Fig.~\ref{fig:ansi-isolation}. For instance,
consider a case where $\I_c(\E) \;=\; \forall
T_i.~\underE{\C{RC}(T_i)}$, a constraint that dictates all
transactions execute under \iso{Read Committed} isolation. Imagine a
sample execution where $\eta_1$'s transaction is not yet committed
when $\eta_2$ is generated. Letting $\eta_1$ be visible to $\eta_2$
violates $\I_c$, whereas not letting it be visible violates $\I_{s}$.
The only way to satisfy both invariants is to rule out all 
executions that interleave the operations of one transaction with the
other, thereby enforcing serializability.  In general, when $\I_s$ conflicts (but is not
inconsistent) with $\I_c$, the only way to enforce both invariant sets
is to restrict concurrency. Clearly, this is unacceptable since it
defeats the very purpose of supporting weak isolation. 
% How then do we enforce weak isolation on a strongly consistent
% machine?

@@

We shall now consider the proof of the example in
Fig.~\ref{fig:motiv-eg-1} in greater detail. We assume an {\sc sc}
store, such as a relational database, whose store-specific trace
invaraint ($\I_s$) was shown previously. Both transactions are run at
{\sc si} isolation, hence $\I_c$ is $\lambda\E.~\underE{\C{SI(Wd1)}}
\conj \underE{\C{SI(Wd1)}}$. As usual, $\I$ is $\I_s \conj \I_c$.

@@

Decorated parts of \C{Wd1} relevant to the discussion are shown in
Fig.~\ref{fig:wd1-decorated}.  \C{Wd2} is similar and not
shown\footnote{Supplementary contains fully decorated program.}.
Temporary name \C{v3} has been introduced to capture the result of the
RHS expression \C{C-a2}.  Assertions implicitly refer to the current
execution ($\E$), just as hoare triples implicitly refer to the
current state.  The context for propositions is also the implicit,
i.e., we write $\psi$ instead of $\underE{\psi}$. The proposition \C{k
$\ge$ a1+a2} remains an invariant, hence elided. The precondition
$P_1$ of \C{Wd1} accounts for the possibility of \C{Wd2} committing
before \C{Wd1}, writing \C{k-a2} to \C{C}. Since neither \C{Wd1} nor
\C{Wd2} are committed at the beginning, the precondition $P$ of the
program is extended with $\neg\committed(\C{Wd1}) \wedge
\neg\committed(\C{Wd2})$, from which $P_1$ follows. The postcondition
$Q_1$ of \C{Wd1} asserts different values for \C{C} depending on
whether or not \C{Wd2} has already committed.  It also asserts that
\C{Wd1} has committed. The postcondition $Q_2$ of \C{Wd2} (not shown)
is similar, but with the roles of \C{Wd1} and \C{Wd2} exchanged. The
postcondition $Q$ follows from $Q_1 \wedge Q_2$, hence valid. It
remains to show that \C{Wd1} satisfies its specification (\C{Wd2} must
also satisfy its spec, but the proof is similar).

First, we focus on the sequential aspect of the proof and show that
the assertions that decorate \C{Wd1} are indeed valid. We consider the
triple associated with the assignment to \C{C} for illustration. The
precondition $\phi_4$ asserts different values for \C{C} and \C{v3}
(temporary alias of RHS expression) depending on whether or not
\C{Wd2} is committed. Inside \C{Wd1}, $\committed(\C{Wd2})$ may mean
that either \C{Wd2} happened before \C{Wd1} (hence can be visible to
all of \C{Wd1}), or that it committed concurrently with \C{Wd1} (hence
cannot be visible to all of \C{Wd2}).  Nonetheless, $\phi_4$ only
considers the case when $\C{Wd2} \visar \C{Wd1}$. As we show below,
this is sufficient to prove the postcondition $\phi_5$, which appears
to make a stronger claim about the value of \C{C}. The proof follows
from the RG rule \rulelabel{RG-Asgn} for assignments, which allows us
to easiliy conclude the following about the execution state after the
assignment:
>>>>>>> 70ae6b11c8f648c323413d6fe8f8fb1c2cba75ae
\begin{smathpar}
\begin{array}{l}
    {\neg\committed(\C{Wd1}) \conj \C{Wd1} \wrstoar \C{C} 
    \conj (\neg\committed(\C{Wd2})} \Rightarrow \C{C = k-a1}) \conj\\
    ({\committed(\C{Wd2})} \Rightarrow \C{Wd2} \wrstoar
    \C{C}) \conj \\
    ({\committed(\C{Wd2})} \wedge {\C{Wd2} \visar \C{Wd1}} 
    \Rightarrow \C{C = k-a2} \wedge \C{C = k-a2-a1})
\end{array}
\end{smathpar}

>>>>

A strongly-consistent {\sc sc} store guarantees a total order on all
operations w.r.t $\visZ$ consistent with their chronological order. A
straightforward $\I_s$ for this store is the \C{SC} property
formalized below:
\begin{smathpar}
\begin{array}{l}
  \C{SC}(\E) \;=\; \forall\eta_1,\eta_2.~\{\eta_1,\eta_2\}
  \subseteq \E.\A \conj \id(\eta_1) < \id(\eta_2) \\
  \hspace*{2in}\Rightarrow \underE{\eta_1 \visar \eta_2}
\end{array}
\end{smathpar}
Unfortunately, $\I_s=\C{SC}$ conflicts with all isolation
specifications of Fig.~\ref{fig:ansi-isolation}. For instance,
consider a case where $\I_c(\E) \;=\; \forall
T_i.~\underE{\C{RC}(T_i)}$, a constraint that dictates all
transactions execute under \iso{Read Committed} isolation. Imagine a
sample execution where $\eta_1$'s transaction is not yet committed
when $\eta_2$ is generated. Letting $\eta_1$ be visible to $\eta_2$
violates $\I_c$, whereas not letting it be visible violates $\I_{s}$.
The only way to satisfy both invariants is to rule out all 
executions that interleave the operations of one transaction with the
other, thereby enforcing serializability.  In general, when $\I_s$ conflicts (but is not
inconsistent) with $\I_c$, the only way to enforce both invariant sets
is to restrict concurrency. Clearly, this is unacceptable since it
defeats the very purpose of supporting weak isolation. 
% How then do we enforce weak isolation on a strongly consistent
% machine?

@@

We now consider the proof of the example in Fig.~\ref{fig:motiv-eg-1}
in greater detail. We assume an {\sc sc} store, such as a relational
database, whose store-specific trace invaraint ($\I_s$) was shown
previously. Both transactions are run at {\sc si} isolation, hence
$\I_c$ is $\lambda\E.~\underE{\C{SI(Wd1)}} \conj
\underE{\C{SI(Wd1)}}$. As usual, $\I$ is $\I_s \conj \I_c$.

@@

Decorated parts of \C{Wd1} relevant to the discussion are shown in
Fig.~\ref{fig:wd1-decorated}.  \C{Wd2} is similar and not
shown\footnote{Supplementary contains fully decorated program.}.
Temporary name \C{v3} has been introduced to capture the result of the
RHS expression \C{C-a1}.  Assertions implicitly refer to the current
execution ($\E$), just as hoare triples implicitly refer to the
current state.  The context for propositions is also the implicit,
i.e., we write $\psi$ instead of $\underE{\psi}$. The proposition \C{k
$\ge$ a1+a2} remains an invariant, hence elided. The precondition
$P_1$ of \C{Wd1} accounts for the possibility of \C{Wd2} committing
before \C{Wd1}, writing \C{k-a2} to \C{C}. Since neither \C{Wd1} nor
\C{Wd2} are committed at the beginning, the precondition $P$ of the
program is extended with $\neg\committed(\C{Wd1}) \wedge
\neg\committed(\C{Wd2})$, from which $P_1$ follows. The postcondition
$Q_1$ of \C{Wd1} asserts different values for \C{C} depending on
whether or not \C{Wd2} has already committed.  It also asserts that
\C{Wd1} has committed. The postcondition $Q_2$ of \C{Wd2} (not shown)
is similar, but with the roles of \C{Wd1} and \C{Wd2} exchanged. The
postcondition $Q$ follows from $Q_1 \wedge Q_2$, hence valid. It
remains to show that \C{Wd1} satisfies its specification (\C{Wd2} must
also satisfy its spec, but the proof is similar).

First, we focus on the sequential aspect of the proof and show that
the assertions that decorate \C{Wd1} are indeed valid. We consider the
triple associated with the assignment to \C{C} for illustration. The
precondition $\phi_4$ asserts different values for \C{C} and \C{v3}
(temporary alias of RHS expression) depending on whether or not
\C{Wd2} is committed. Inside \C{Wd1}, $\committed(\C{Wd2})$ may mean
that either \C{Wd2} happened before \C{Wd1} (hence can be visible to
all of \C{Wd1}), or that it committed concurrently with \C{Wd1} (hence
cannot be visible to all of \C{Wd2}).  Nonetheless, $\phi_4$ only
considers the case when $\C{Wd2} \visar \C{Wd1}$. As we show below,
this is sufficient to prove the postcondition $\phi_5$, which appears
to make a stronger claim about the value of \C{C}. The proof follows
from the RG rule \rulelabel{RG-Asgn} for assignments, which allows us
to easily conclude the following about the execution state after the
assignment:

\begin{smathpar}
\begin{array}{l}
    {\neg\committed(\C{Wd1}) \conj \C{Wd1} \wrstoar \C{C} 
    \conj (\neg\committed(\C{Wd2})} \Rightarrow \C{C = k-a1}) \conj\\
    ({\committed(\C{Wd2})} \Rightarrow \C{Wd2} \wrstoar
    \C{C}) \conj \\
    ({\committed(\C{Wd2})} \wedge {\C{Wd2} \visar \C{Wd1}} 
    \Rightarrow \C{C = k-a2} \wedge \C{C = k-a2-a1})
\end{array}
\end{smathpar}
